---
title: "Exploratory Report"
author: "Patrick Cheng, Erica Gordy, William Siauw, Joel Yang"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

```{r setup, include=FALSE}
source("analysis.R")
library(tidyverse)
library(dplyr)
library(ggplot2)
library("maps")
```

## Section 2.1 #1

The datasets detail raw statistics such as income in cities and counties across the United States, crime statistics in Washington, and incarcerated populations by state.

## Section 2.1 #2

Most of these datasets come from [Kaggle](https://www.kaggle.com/), an online data science environment where numerous data sets are published for free use. The exception is our “Criminal Justice Data Book” set, which comes from [Data.gov](https://www.data.gov/), a government hosted website.

## Section 2.1 #3

The crime dataset looks something like this:
```{r sample, echo=F}
head(crime_2016_total)
```

The income dataset looks something like this:
```{r, echo=F}
head(modified_income_df)
```

## Section 2.1 #4

Our modified income dataframe is relatively straightforward, giving secondary statistics such as the mean, median and standard deviation of
each respective county in Washington. sum_w refers to the number of houses sampled per county.

Our crime in 2016 dataframe is a bit more confusing, but it becomes more readable once the abbreviations are understood.
- PDP refers to prison admits,
- JDP refers to jail admits,
- and POP refers to total population (including the incarcerated).


## Section 2.2 #1 

For the crime data frame, we took the mean values for each age group. Then we went county by county gathering mean, median, min, and max for both Male and Female. For this we used the POP_F_TOTAL and POP_M_TOTAL column data. 

For the income data frame, we began by looking at national data. We also added a column so that we could see the county and state names together. We looked at mean, median, min, and max data nationally and then focused on WA state, repeating the same stats for WA counties. 

Below are some of the most relevant statistics from the data sets. 
Crime Data in WA state:
Mean Female Incarcerated
```{r, echo=F}
print(mean_F_total)
```
Mean Male Incarcerated
```{r, echo=F}
print(mean_M_total)
```
County with the Most Female Incarcerated
```{r, echo=F}
print(max_county_name_F)
```
County with the Most Male Incarcerated
```{r, echo=F}
print(max_county_name_M)
```
County with the Least Female Incarcerated
```{r, echo=F}
print(min_county_name_F)
```
County with the Least Male Incarcerated
```{r, echo=F}
print(min_county_name_M)
```

Income Data: 
Mean National Income:
```{r, echo=F}
print(mean_national_income)
```
Washington State Statistics:
```{r, echo=F}
print(WA_state_income)
```
Washington State Statistics By County:
```{r, echo=F}
print(WA_counties_income)
```

## Section 2.2 #2

```{r, echo=F}
plot(jail_wa_plot)
```

```{r, echo=F}
plot(jail_density_plot)
```

```{r, echo=F}
plot(top10_jailrate_plot)
```


```{r, echo=F}
plot(income_wa_plot)

```
```{r, echo=F}
plot(income_density_plot)
```

```{r, echo=F}
plot(top_incomes_plot)
```

### Outliers

In this income graph, there are a few counties with no data. This missing data could be due to a plethora of reasons, such as maybe a low population in said county, or simply not enough information. Additionally, Douglas county is missing daily average jail population data. We aren't too sure what might result in this, but possibly poor reporting or large swathes of missing information.


## Section 2.3

### Are people of different economic status punished by law differently in the state of Washingtion? 








### How is the relationship between the rate of people getting arrested and the rate of people being sentenced to prison affected by economic status in the state of Washington?

The rate of people getting arrested is the number of total people arrested in a year divided by total population of the county. The rate of prison sentence is the number of total prison admission in a year divided by total population of the county.

The relationship between the arrest rate and prison sentence rate in a certain county is calculated by dividing the total number of prison admission in a year by total number of people arrested in that year. Then I compare the result to the median income of the county.

```{r, Echo = F }
arrest_vs_prison_data %>% 
  select(county, prison_to_arrest_rate, income_median)

correlation <- cor(arrest_vs_prison_data$prison_to_arrest_rate,
                   arrest_vs_prison_data$income_median)
```

The correlation coefficient between prison-arrest rate and income median is `r correlation`

```{r, echo=F}
arrest_vs_prison_by_median_plot
```

#### Evaluation

?????????????????????
Outlier

### How does the economic status of residents in a given county correlate with the incarceration rate in the state of Washington?

For this question, we wanted to examine if there were any correlations between a county's mean income and it's rate of incarceration.

In order to answer this question, we simply compared the two maps exploring economic status in Washington to the map exploring incarceration across Washington.

### How does the type of crime committed relate to the geographic location of the crime and the economic status of the area in the state of Washington?



